### 2015-06-05 LOI/inc2.tex
Releasing lock

% \item What is the new technical idea; why can we succeed now?
%  Integration of proven components into an architecure that solves
%  problems. New research will resolve remaining technical hurdles.

@@

%
Our objective here is to reduce the time to knowledge, an end-to-end metric
relevant to the scientific discovery process. Beyond the traditional high
volume I/O pattern of checkpoint/restart, we will address the challenges
posed by additional data access patterns in the knowledge gathering
process. With a deeper insight into the scientific process we will encode
and utilize errors and accuracy as an optimization parameter. For example, 
%
% We will explore access to
% large data sets as part of this process, w

% which means that we will
% also support additional access modes, beyond the traditional
% Checkpoint/Restart methods, and understand that all data has errors
% associated with it and we will ensure that these become encoded to allow for
% additional optimizations. 
For example, all scientific simulations contain
approximations, and all measurements from observations and experiments also
contain errors. These errors can help guide how information can be presented
to users, by allowing us to ask for information within a given accuracy
bound, and also allow us to offer a mode for re-computation vs. data storage
and reterival.

@@

\underline{\textbf{Key Technical Approach:}}
%We are fundamentally trying to address the fundmantal questions : 1)  How can we 
%place and mangage massive scienfici data across all of the tiers of the storage and memory
%hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
%to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
%that the placement of information can be agreed upon from what the application requires and what the
%system resources are available.

Our approach will allow users to ``plug-in'' techniques to classify data, not as bytes but as motifs, where
we can understand relationships between objects (into data models) and relationships of data in variables.
This will allow us to adapt data from a variable into the various storage tiers. Rather than focus on simple data
compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
reduced information on the lower tiers.

We wish to support additional data access modes as well. First, within a given
accuracy bounds, offer a mode for re-computation from data stored in a ``fast''
tier to reduce data latencies. Second, exploratory analysis operations
frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
Third, to ensure available storage for subsequent operations, we will offer
automatic data migration based on user annotations for required data lifetimes
using monitoring and learning techniques. Unlike existing approaches, this will
be tempered both by the user annotations and through learned access patterns.
While past access patterns may not indicate future access because the
simulation run purpose may have changed, we are focused on scalability where
runs are subsequently larger as the simulation prepares for a capability run.
By learning from the output and access patterns during this run sequence, we
can accurate decide how to place and organize data for the critical capability
runs. Fourth, we anticipate storing multiple data copies, each compressed in
different ways according to the underlying media, some of these copies will
disappear based on storage pressures, but data persistence will be maintained
according to user specifications. Assuming a relatively low latency cache layer
before a tape system, we can offer exploratory data access reserving pulling
data from tape to just the data required. This will save scientists time and
make data stored on tape usable without long delays.

@@

The success of this project will  provide insights into how to build middleware and storage layers which can interact
well with each other, and take user-provided hints. Today data is reduced by application scientist who have 
limited information on what the storage layer can provide. They often make compromises based on this limited knowledge
and either tune their output for writing or reading performance. This data then gets moved to other locations, and much of
the tuning is lost when the data is read back during their post processing. Furthermore, there is a limited set of operations
which users will be able to stage to other staging nodes for real-time-reduction and visualization.

We will identify,   through concrete application evaluation, the requirements for highly 
usable and scalable middleware and storage layers

>>>>

% \item What is the new technical idea; why can we succeed now?
%  Integration of proven components into an architecture that solves
%  problems. New research will resolve remaining technical hurdles.

@@

%
Our objective here is to reduce the time to knowledge, an end-to-end metric
relevant to the scientific discovery process. Beyond the traditional high
volume I/O pattern of checkpoint/restart, we will address the challenges
posed by other essential data access patterns in the knowledge gathering
process. Through a deeper insight into the scientific process we will encode
and utilize accuracy and errors as optimization parameters. 
%
% We will explore access to
% large data sets as part of this process, w

% which means that we will
% also support additional access modes, beyond the traditional
% Checkpoint/Restart methods, and understand that all data has errors
% associated with it and we will ensure that these become encoded to allow for
% additional optimizations. 
For example, scientific simulations contain approximations, as do measurements 
from observations and experiments, and depending on the goal, these approximations 
are acceptable. We can leverage this observation to optimize the presentation of data
to the user and to implement various tradeoffs. For example, users can ask for information 
within a given accuracy bound, allowing us to offer a mode for re-computation vs. data storage
and reterival.

@@

\underline{\textbf{Key Technical Approach:}}
%We are fundamentally trying to address the fundmantal questions : 1)  How can we 
%place and mangage massive scienfici data across all of the tiers of the storage and memory
%hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
%to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
%that the placement of information can be agreed upon from what the application requires and what the
%system resources are available. 
Our overall technical approach is based on an application-aware runtime realization of tradeoffs in 
data-representation, data placement and data access. For example, we will allow users to ``plug-in'' 
techniques to classify data, not as bytes but as motifs, allowing us to understand relationships 
between objects (based on data models), the relationships of data and variables, as well as data
access and transport patterns.  This will allow us 
to adaptively map data from a variable onto the various storage tiers. Rather than focus on simple data
compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
reduced information in the lower storage tiers.

We will support additional data access modes as well. First, within a given
accuracy bounds, offer a mode for re-computation from data stored in a ``fast''
tier to reduce data latencies. Second, exploratory analysis operations
frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.

Third, to ensure available storage for subsequent operations, we will offer
automatic data migration based on user annotations for required data lifetimes
using monitoring and learning techniques. Unlike existing approaches, this will
be tempered both by the user annotations and through learned access patterns.
While past access patterns may not indicate future access because the
simulation run purpose may have changed, we are focused on scalability where
runs are subsequently larger as the simulation prepares for a capability run.
By learning from the output and access patterns during this run sequence, we
can accurate decide how to place and organize data for the critical capability
runs. Fourth, we anticipate storing multiple data copies, each compressed in
different ways according to the underlying media, some of these copies will
disappear based on storage pressures, but data persistence will be maintained
according to user specifications. Assuming a relatively low latency cache layer
before a tape system, we can offer exploratory data access reserving pulling
data from tape to just the data required. This will save scientists time and
make data stored on tape usable without long delays.

@@

The success of this project will  provide insights into how to build autonomic middleware and storage layers which can interact
well with each other, and take user-provided hints. Today, data is reduced by application scientist who have 
limited information on what the storage layer can provide. They often make compromises based on this limited knowledge
and either tune their output for writing or reading performance. This data then gets moved to other locations, and much of
the tuning is lost when the data is read back during their post processing. Furthermore, there is a limited set of operations
which users will be able to stage to other staging nodes for real-time-reduction and visualization.

%We will identify,  through concrete application evaluation, the requirements for highly usable and scalable middleware and storage layers
